A molecular taxonomy of tumors independent of tissue-of-origin

### 2020-08-21 code/survival/survfig.Rmd

#### Figure 7
```{r, survfig, echo=FALSE, warning=FALSE, error=TRUE,message=FALSE,fig.height=9, fig.width=9}

@@

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 8
```{r, message = FALSE, fig.width = 9}
mylayout = '
ABB
ABB
#BB
'

((pcanc+mytheme) + (pclass+mytheme))+plot_layout(design=mylayout) +plot_annotation(tag_levels = "A")
```

>>>>

```{r, survfig, echo=FALSE, warning=FALSE, error=TRUE,message=FALSE,fig.height=9, fig.width=9}

@@

### 2020-08-21 code/table1.Rmd

```{r table-1-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/moderate_enrich_prob.Rda") ## load data
load("../data/gene2_level2_lookup.Rda")
load("../data/gene_data.Rda")
load("../data/gdc_gene_exp_adj.Rda")
load("../data/pathname2level2.Rda")

id_to_names <- gene2level2_lookup %>% rownames_to_column(var = "ensembl_gene_id") %>% ## convert ensembl to gene name
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>%
  remove_rownames() %>% column_to_rownames(var = "external_gene_name") %>%
  dplyr::select(-ensembl_gene_id) %>%
  t() %>% as.data.frame() %>% rownames_to_column(var = "PATH_ID") %>% ## convert path_id to path_name
  left_join(., pathname2level2, by = "PATH_ID") %>%
  remove_rownames() %>% column_to_rownames(var = "PATH_NAME") %>%
  dplyr::select(-PATH_ID) %>%
  t() %>% as.data.frame()

gene_mutation <- gdc_gene_exp_adj %>% rownames_to_column(var = "ensembl_gene_id") %>% ## convert ensembl to gene name
  left_join(., gene_data[,c("ensembl_gene_id", "external_gene_name")], by = "ensembl_gene_id") %>%
  remove_rownames() %>% column_to_rownames(var = "external_gene_name") %>%
  dplyr::select(-ensembl_gene_id)

## parameters ##
enrich_cutoff <- 0.30
howmanypaths <- 5 ## top _ pathways; how many top pathways do you want to see
show_perc <- TRUE
show_n <- TRUE
mutated_gene_position <- 1
################

clust_spec_enrich <- moderate_enrich_prob %>% filter(signif <= 2500 & difference >= enrich_cutoff) ## signif is 95% CI with a specified cutoff

clust_spec_enrich_top5 <- list()
for(n in unique(clust_spec_enrich$clust)) {
  message(n)
  clust_spec_enrich_top5[[n]] <- clust_spec_enrich %>% filter(clust == n) %>% ## filter for specific cluster
    arrange(desc(difference)) %>% ## arrange enrichment score in decreasing order 
    dplyr::select(path, clust) %>% .[1:5,]
}
clust_spec_enrich_top5 <- clust_spec_enrich_top5 %>% do.call(rbind, .) %>% as.data.frame() ## combine list from for loop
rm(clust_spec_enrich)

## finding the top mutated gene for each cluster in each pathway for the top 5 pathways 
type <- unique(tumor_clusters$project_code) ## type of cancer type in data
cluster_spec_enrich_table <- list() ## list for forloop

for(d in unique(clust_spec_enrich_top5$clust)) {
  message(d)
  clust_toppath <- clust_spec_enrich_top5 %>% filter(clust == d) %>% .$path %>% as.character() ## get the top pathway names for spec class
  class_spec <- tumor_clusters %>% dplyr::select(sample_id, clust_knn, project_code) %>% filter(clust_knn %in% d) ## get the smaple ids and cancer type for spec class
  top_genes_in_path_cancer_type <- list()
  for(n in clust_toppath) { ## for each of the top paths do this
    message(n)
    gene_in_path <- id_to_names %>% .[,which(names(.) == n)] %>% as.data.frame() %>% ## select pathway
    'rownames<-' (rownames(id_to_names)) %>% 
    filter(. == 1) %>% ## filter for genes that are in that pathway
    rownames(.)
    top_gene_in_path <- list()
    for(t in type) { ## looping through cancer type within class
      message(t)
      class_spec <- tumor_clusters %>% dplyr::select(sample_id, clust_knn, project_code) %>% filter(clust_knn == d) ## sample_id for class
      if(t %in% class_spec$project_code) { ## if cancer type is in class then run this 
        class_type_spec <- class_spec %>% filter(project_code == t) %>% column_to_rownames(var = "sample_id") %>% rownames() ## get cancer type specific tumor classes
      if(length(class_type_spec) == 1) { ## if you only have 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% rownames_to_column(var = "gene") %>% .[.$gene %in% gene_in_path, names(.) %in% c("gene",  class_type_spec)] %>% remove_rownames() %>% column_to_rownames(var = "gene") %>% 'colnames<-' ("obs_count") %>%
          rownames_to_column(var = "gene") %>%
          mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
         arrange(desc(obs_count))
      }
      if(length(class_type_spec) > 1) { ## if you have more than 1 sample_id in the class
        top_mut_gene <- gene_mutation %>% .[rownames(.) %in% gene_in_path, names(.) %in% class_type_spec] %>% 
        rowSums() %>% as.data.frame() %>% 'colnames<-' ("obs_count") %>%
        rownames_to_column(var = "gene") %>%
        mutate(n_sample = length(class_type_spec),
               perc_mut = round((obs_count / length(class_type_spec)) * 100, digits = 1)) %>% 
         arrange(desc(obs_count))
      }
      if(show_perc == TRUE & show_n == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)")  
      }
      if(show_perc == TRUE & show_n == TRUE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position], " (", top_mut_gene$perc_mut[mutated_gene_position],"%)", "    n = ", length(class_type_spec))  
      }
      if(show_perc == FALSE) {
         top_gene_in_path[[t]] <- paste0(top_mut_gene$gene[mutated_gene_position])  
      }
      }
      if(!(t %in% class_spec$project_code)){ ## if cancer type is not in class then insert NA
        top_gene_in_path[[t]] <- "NA" 
      }
    }
    top_gene_in_path <- top_gene_in_path %>% do.call(rbind, .) %>% as.data.frame() %>% 'colnames<-' (n)
    clust_row <- data.frame(Class = d, row.names = n) %>% t() ## creating a row for cluster membership
    top_genes_in_path_cancer_type[[n]] <- rbind(clust_row, top_gene_in_path)
  }
  cluster_spec_enrich_table[[n]] <- top_genes_in_path_cancer_type %>% do.call(cbind, .) %>% as.data.frame() %>% 
  t() %>% as.data.frame() %>% rownames_to_column(var = "Pathway")
  }

cluster_spec_enrich_table <- cluster_spec_enrich_table %>% do.call(rbind, .) %>% as.data.frame()
save(cluster_spec_enrich_table, file = "../data/cluster_spec_enrich_table.Rda")
rm(list = ls(all.names = T))
gc()
```

@@

## Selecting the top 10 cancers with the largest n for table 1
n_tumor <- as.data.frame(table(tumor_clusters$project_code))
n_tumor <- n_tumor %>% arrange(desc(Freq)) %>% .[1:10, "Var1"] %>% as.character() %>% .[order(.)]

cluster_spec_enrich_table %>% .[, c("Pathway", "Class", n_tumor)] %>% knitr::kable(row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = T)
```

>>>>

```{r table-1-data-generation, eval = FALSE}
tumor_clusters <- fread("../data/umap_3d_coors.tsv") ## load in data
load("../data/gdc_reactome_path_analysis.Rda")

tumor_clusters <- tumor_clusters %>% dplyr::select(sample_id, ajcc_metastasis_pathologic_pm, clust_knn)

m0_clust <- tumor_clusters[tumor_clusters$ajcc_metastasis_pathologic_pm == "M0",] ## filter for nonmetastatic samples
m1_clust <- tumor_clusters[tumor_clusters$ajcc_metastasis_pathologic_pm == "M1",] ## filter for metastatic samples

prepare_rbeta_data_p1 <- function(data, path_analysis) {
  n_tumor_in_group <- as.data.frame(table(data$ajcc_metastasis_pathologic_pm)) ## Getting the total number of tumor samples in each group
  path_analysis <- path_analysis[,names(path_analysis) %in% data$sample_id] ## Filter only group specific cancers
  data <- data[match(names(path_analysis), data$sample_id),] ## Match tumor sample ids
  table(names(path_analysis) == data$sample_id) ## good they match now
  ## Now we are going to build group specific cluster data
  n_tumor_in_clust <- as.data.frame(table(data$clust_knn)) ## getting total n in each cluster
  names(n_tumor_in_clust) <- c("clust", "tot_n_clust") ## rename colnames
  clust_group <- as.character(n_tumor_in_clust$clust) ## Getting the cluster numbers
  cluster_path_count <- list() ## Creating a list for the forloop
  for(n in clust_group) {
    print(n)
    df <- subset(data, clust_knn == n) ## subset cluster number
    df <- df$sample_id ## getting tumor sample id
    df1 <- path_analysis[,names(path_analysis) %in% df] ## getting the tumor samples from path analysis
    df1 <- as.data.frame(rowSums(df1))
    names(df1) <- n
    cluster_path_count[[n]] <- df1
  }
  cluster_path_count <- do.call(cbind, cluster_path_count) ## Combine the list together
  rbeta_path_df <- cluster_path_count %>% rownames_to_column(var="path") 
  rbeta_path_df <- rbeta_path_df %>% gather(key = "clust", value = "mutation", -path) ## Wide to long format
  global_path_mut <- as.data.frame(rowSums(cluster_path_count))
  global_path_mut <- global_path_mut[rep(1:ncol(global_path_mut), each = max(as.numeric(as.character(rbeta_path_df$clust))))] ## Repeat column 12 times
  names(global_path_mut) <- as.character(clust_group) ## Replace names with cluster numbers
  global_path_mut <- global_path_mut %>% rownames_to_column(var="path") %>% gather( key = "clust", value = "global", -path) ## Wide to long format
  rbeta_path_df <- left_join(rbeta_path_df, global_path_mut, by = c("path", "clust")) ## left_join rbetadf and global
  rbeta_path_df <- left_join(rbeta_path_df, n_tumor_in_clust, by = "clust")
  rbeta_path_df$global_sample <- as.numeric(sum(n_tumor_in_clust$tot_n_clust)) ## Getting total tumor sample number
  names(rbeta_path_df) <- c("path", "clust", "clust_mut", "global_mut", "tot_n_clust", "global_sample")
  rbeta_path_df$tot_n_clust <- as.numeric(as.character(rbeta_path_df$tot_n_clust)) ## convert tot_n_clust numeric
  return(rbeta_path_df)
}

m0_rbeta <- prepare_rbeta_data_p1(data = m0_clust, path_analysis = gdc_reactome_path_analysis)
save(m0_rbeta, file = "../data/m0_rbeta.Rda")

## repeat same function above with metastasis sample
m1_rbeta <- prepare_rbeta_data_p1(data = m1_clust, path_analysis = gdc_reactome_path_analysis)
save(m1_rbeta, file = "../data/m1_rbeta.Rda")

## function to compiling data for rbeta analysis
prepare_rbeta_data_p2 <- function(data) {
  enrich_prob <- data.frame(path = data$path,
                               clust = data$clust,
                               fg_mut = data$clust_mut,
                               fg_not_mut = data$tot_n_clust - data$clust_mut,
                               bg_mut = data$global_mut,
                               bg_not_mut = data$global_sample - data$global_mut)
fg_mut_rate <- list()
bg_mut_rate <- list()
for(n in unique(as.character(enrich_prob$clust))) {
  print(n)
  df <- subset(enrich_prob, clust == n)
  fg_denominator <- nrow(df) * (df$fg_mut[1] + df$fg_not_mut[1]) ## n_path * tot_n_clust_x
  fg_numerator <- sum(df$fg_mut)
  bg_denominator <- nrow(df) * (df$bg_mut[1] + df$bg_not_mut[1])
  bg_numerator <- sum(df$bg_mut)
  fg_mut_rate[[n]] <- fg_numerator / fg_denominator
  bg_mut_rate[[n]] <- bg_numerator / bg_denominator
}

fg_mut_rate <- as.data.frame(do.call(rbind, fg_mut_rate)) ## rbind the mut rates from list
bg_mut_rate <- as.data.frame(do.call(rbind, bg_mut_rate)) ## rbind the mut rates from list
enrich_prob_df <- enrich_prob
enrich_prob_df$fg_mut_rate <- rep(x = fg_mut_rate$V1, each = length(unique(enrich_prob_df$path)))
enrich_prob_df$bg_mut_rate <- rep(x = bg_mut_rate$V1, each = length(unique(enrich_prob_df$path)))
enrich_prob_df$rel_odds <- enrich_prob_df$fg_mut_rate / enrich_prob_df$bg_mut_rate
enrich_prob_df$denominator <- (enrich_prob_df$bg_mut * enrich_prob_df$rel_odds) +   (enrich_prob_df$bg_not_mut / enrich_prob_df$rel_odds)
enrich_prob_df$x_adj <- round((enrich_prob_df$fg_mut + enrich_prob_df$fg_not_mut) * (enrich_prob_df$bg_mut * enrich_prob_df$rel_odds) / enrich_prob_df$denominator)
enrich_prob_df$y_adj <- (enrich_prob_df$fg_mut + enrich_prob_df$fg_not_mut) -   enrich_prob_df$x_adj
return(enrich_prob_df)
}

m0_rbeta <- prepare_rbeta_data_p2(m0_rbeta)                               
m1_rbeta <- prepare_rbeta_data_p2(m1_rbeta)

save(m0_rbeta, m1_rbeta, file = "../data/m0_m1_rbeta.Rda")
rm(list = ls(all.names = T)) ## cleaning up data
gc()
load("../data/m0_m1_rbeta.Rda")
## Metastatic (foreground) - Nonmetastatic (background) Rbeta
moderate_fg <- apply(m1_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})
moderate_bg <- apply(m0_rbeta[,c("fg_mut", "fg_not_mut")], MARGIN = 1, function(x) {
  rbeta(n = 1e5, shape1=x["fg_mut"], shape2=x["fg_not_mut"])
})

metastasis_cluster_moderate <- moderate_fg - moderate_bg  ## Fg - Bg ... to see how many times the Fg is larger than the Bg
metastasis_cluster_moderate <- as.data.frame(metastasis_cluster_moderate)

## ModerateEnrichment probability 
prob <- moderate_fg > moderate_bg ## boolean data
mean_prob <- apply(prob, MARGIN = 2, function(x) {
  mean(x)
})
## Moderate Difference
CI <- 0.025

diff_fg_bg <- moderate_fg - moderate_bg ## fg - bg
rm(moderate_fg, moderate_bg, prob)
gc()
quant_diff <- t(apply(diff_fg_bg, MARGIN = 2, function(x) {
  quantile(x, c(CI, 0.5, 1-CI)) ## 95% CI
}))

sig_fg_bg <- as.data.frame(apply(diff_fg_bg, MARGIN = 2, function(x) {
  sum(x < 0 ) ## out of 1e5 simulations, how many times is fg-bg less than 0
}))

m1_minus_m0_per_cluster_rbeta <- data.frame(path = m1_rbeta$path,
                                   clust = m1_rbeta$clust,
                                   probability = mean_prob,
                                   signif = sig_fg_bg[,1],
                                   difference = quant_diff[,2],
                                   lower = quant_diff[,1],
                                   upper = quant_diff[,3])
save(m1_minus_m0_per_cluster_rbeta, file = "../data/m1_minus_m0_per_cluster_rbeta.Rda")
rm(list = ls(all.names = T))
gc()
```

@@

```{r table-2, eval = FALSE}
## this is for html only
load("../data/m1_minus_m0_per_cluster_rbeta.Rda")

table2 <- m1_minus_m0_per_cluster_rbeta %>% filter(difference >= 0.20 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

table2$log10pvalue <- round(-log10((table2$Signif + 1) / 1E5), digits = 2) ## Getting p value ## adding 1 for pseudo count.. 1E5 is the n simulations performed
table2$Enrichment_Score <- round(table2$Enrichment_Score, digits = 2)
table2$Signif <- NULL ## we don't need this anymore

table2 %>% knitr::kable() %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"), full_width = FALSE, font_size = 12) %>%
  column_spec(1, bold = T, color = "black")
```

```{r table2-docx-friendly}
## docx friendly
load("../data/m1_minus_m0_per_cluster_rbeta.Rda")

table2 <- m1_minus_m0_per_cluster_rbeta %>% filter(difference >= 0.30 & signif <= 2500) %>% ## filter effect size >= 0.20 and enrichment at a 95% CI
  dplyr::select(path, clust, signif, difference) %>% 'colnames<-' (c("Pathway", "Cluster", "Signif", "Enrichment_Score")) ## select relevant column and rename col

table2$log10pvalue <- round(-log10((table2$Signif + 1) / 1E5), digits = 2) ## Getting p value ## adding 1 for pseudo count.. 1E5 is the n simulations performed
table2$Enrichment_Score <- round(table2$Enrichment_Score, digits = 2)
table2$Signif <- NULL ## we don't need this anymore

table2 <-table2 %>% flextable() %>% 
   set_caption(caption = "") %>%
   align(part = "all") %>% 
   width(., width = c(4,0.75,1,1)) %>% 
   fontsize(size = 10, part = "body")

knitr::knit_print(table2)
```

### 2020-08-21 text/abstract.Rmd

Cancer is a disease of cellular metabolism and other basic biochemical processes
common to all cells. However, molecular signatures remain largely
indistinguishable from tissue-of-origin, and general, unifying principles remain
obscure. To address this challenge, we reduced mutation data to disruptions in a
select set of pathways relevant to basic cell biology, from DNA replication to
cellular communication. Using dimension reduction techniques, we assigned tumor
samples into ten classes distinct from their tissues-of-origin and largely
free of bias from mutational burden or clinical stage. We showed that these
tumor classes vary in prognosis by modeling relative risk of death in different
classes and cancer types. We identified class-specific mutations in different
tissues, demonstrating that tissue-specific signatures contribute to common
cellular phenotypes. Moreover, driver gene mutations involved in replication fidelity and genome
stability, such as BRCA1 and Rb1, are equally distributed among our classes,
contrary to the expectation that such genes are avatars of molecular
subtype. We also investigated pathways that distinguish between metastatic and
non-metastatic tumors, and found that most meaningful differences are
class-specific. Many metastastasis enriched pathways from one class of
tumors are class-specific pathways from another class, suggesting phenotypic
convergence, and potentially highlighting key pathways to focus on
for prevention of class-specific relapse. In summary, our
observations suggest that some genes act as drivers of genome instability in all
classes, while different sets of class-specific genes confer increased fitness
in a combinatorial fashion, eventually converging on one or more end-stage
disease phenotypes. Thus, we present a coherent view of global tumor biology,
and explain how common cellular dysfunction might arise from tissue-specific
mutations.

>>>>

Cancer is a complex disease involving dysfunctional cellular metabolism, other
basic biochemical processes, and the microenvironment. However, in spite of some
generally agreed upon unifying principles [@weinberg2000; @hallmarks_II],
molecular signatures remain largely indistinguishable from tissue-of-origin,
thus obscuring opportunities for novel mechanistic insights and potential
treatment vulnerabilities that hold true across different cancer types. To
address this challenge, we reduced mutation data to disruptions in a select set
of pathways relevant to basic cell biology, from DNA replication to cellular
communication. Using dimensionality reduction techniques, we assigned tumor
samples into ten clusters distinct from their tissues-of-origin and largely free
of bias from mutational burden or clinical stage. We showed that these tumor
classes vary in prognosis by modeling relative risk of death in different
classes and cancer types. We identify cluster-specific mutations in different
tissues, demonstrating that tissue-specific signatures contribute to common
cellular phenotypes. Moreover, driver gene mutations involved in replication
fidelity and genome stability, such as _BRCA1_ and _Rb1_, are equally distributed
among our clusters, contrary to the expectation that such genes are avatars of
molecular subtype. We also investigated pathways that distinguish between
metastatic and non-metastatic tumors, and found that most meaningful differences
are cluster-specific. Some metastasis enriched pathways from one cluster are
cluster-specific pathways from non-metastatic tumors of another cluster,
suggesting phenotypic convergence, and potentially highlighting key pathways to
focus on for prevention of cluster-specific relapse. Taken as a whole, our
observations suggest that common driver genes combine with tissue-specific
disruptions in complementary sets of tumor-promoting pathways to produce a
limited number of distinct molecular phenotypes. Eventually these signatures
converge on one or more end-stage disease phenotypes. Thus, we present a
coherent view of global tumor biology, and explain how common cellular
dysfunction might arise from tissue-specific mutations.

### 2020-08-21 text/survival.Rmd

If our molecular pathway disruption classes represent distinct biological states
apart from tissue of origin, they may have different prognoses within cancer
types or across all cancers. Kaplan-Meier plots reveal large differences between
cancer types as expected, and some modest differences between our classes (not
shown). These analyses are profoundly limited by confounding factors of age and
stage at diagnosis, sex, ethnicity, and tissue-specific disease progression, to
name the obvious ones. To explore these ideas, we tested models of survival
using public longevity data from the CDC, accounting for age, gender and
ethnicity. Our model assumes that at birth each person has a small initial
probability per unit time to die, and that this probability grows exponentially
in time at some baseline rate that matches the observed CDC data. The rate of
growth in risk is assumed to jump to a higher level at diagnosis, which we take
as a proxy for onset. This represents an accelerated aging type model. To fit
the model, we constrain the baseline risks using the available CDC life tables.
We then constrain the cancer specific multiplier using the longevity data within
the cancer tumor dataset. Our model estimates the effect of cancer type and
class-specific cancer effects independently, resulting in a cancer and class
specific estimate of the aging rate multiplier.

@@

@@

>>>>

---
output:
  word_document: default
  html_document: default
---
If our molecular pathway disruption classes represent distinct biological states
apart from tissue of origin, they may have different prognoses within cancer
types or across all cancers. Kaplan-Meier plots reveal large differences between
cancer types as expected, and some modest differences between our classes (not
shown). These analyses are profoundly limited by confounding factors of age and
stage at diagnosis, sex, ethnicity, and tissue-specific disease progression, to
name the obvious ones. To explore these ideas, we tested models of survival
using public longevity data from the CDC, accounting for age, gender and
ethnicity. Our model assumes that at birth each person has a small initial
probability per unit time to die, and that this probability grows exponentially
in time at some baseline rate that matches the observed CDC data. The rate of
growth in risk is assumed to jump to a higher level at diagnosis, which we take
as a proxy for onset. This represents an accelerated aging type model. To fit
the model, we constrain the baseline risks using the available CDC life tables.
We then constrain the cancer specific multiplier using the longevity data within
the cancer tumor dataset. Our model estimates the effect of cancer type and
class-specific cancer effects independently, resulting in a cancer and class
specific estimate of the aging rate multiplier.

@@

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 7
```{r child='../code/survival/survfig.Rmd'}
```

**Figure 7: Comparison of different cancers, simulated and actual results
results.** A) To understand the risk model, we plot survival of four simulated
cohorts. Compare survival of a group consisting of randomly selected 30 year old
women (light blue) to an equal sized group of 30 year old women with cancer, k =
2 (light orange). To understand how the mix of ages alters the survival curve,
we compare a randomly selected group of people with ages uniformly distributed
from 30 to 70 (dark blue) and a randomly selected group with uniformly
distributed ages and cancer k = 2 (dark orange). B,C,D,E) We compare a randomly
selected group with correct age distribution for the given cancer (blue) to the
model predicted survival for the given average k (red) and the actual survival
for the patients in our dataset (CHOOSE COLOR HERE, NOT GREEN). As can be seen,
Stomach Cancer (STAD), Glioblastoma (GBM), and Pancreatic Cancer (PAAD) are all
very deadly. The modeled results fit the general trend for the actual patients.
Thyroid Cancer (THCA) on the other hand has very little if any effect on life
expectancy, in fact there may be a slight benefit which may be associated with a
bias in diagnosis towards patients who are more health conscious, or higher
socioeconomic status, with more access to care compared to random population
members.

```{r, results='asis'}
cat("\\newpage")
```

@@

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 8
```{r fig.width=3, message=FALSE, warning=FALSE}
pcanc+mytheme
```

**Figure 8 Tissue specific effects** Effects are compared between cancer type by
showing the probability density for the posterior value of the $k_{tis}$
multiplier. These values reflect the general deadliness of the cancer type
across all clusters. GBM, PAAD, and SKCM stand out as having values in the 10 to
20 range with a wide degree of uncertainty. Other cancers such as Prostate
(PRAD) and Thyroid (THCA) show values down near or even below 1 suggesting that
these populations may have healthier lifestyles and more access to care. Most
cancer types fall in the range 2-7 as might be expected.

```{r, results='asis'}
cat("\\newpage")
```

#### Figure 9

```{r child='../code/figure9.Rmd'}
```

**Figure 9 Class specific $k_{class}$ values shown for each cancer type.** The
overall $k$ value is the product of $k_{tis} \times k_{class}$. The class
specific value shows the relative aggressiveness of that particular class within
the tissue. For example class 1, 2, and 6 Glioblastoma (GBM) are apparently much
less deadly than other types of GBM. A similar situation can be seen with
Pancreatic Cancer (PAAD) where 1, 6, and 8 are less deadly compared to classes
7, 9, 10. For some cancer types there are few data points and little ability to
estimate values with precision (DLBC and KICH for example).

```{r, results='asis'}
cat("\\newpage")
```
